---
title: "Linear Regression Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linear Regression Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(Lab4)
library(ggplot2)
library(graphics)
library(methods)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette describes the function **`linreg_f`** to create a linear regression model and explains how to use the implemented methods of the RC class **`linreg`** to analyze the model. 

## Function linreg_f

The function linreg_f has two arguments `formula` and `data`. The argument `formula` is a `formula` object in R and `data` is a sample dataset e.g. `iris`, `cars`. The function returns an object of RC class `linreg`. 

In the example shown below, class object `linreg_mod` is a linear regression model built based on the `formula` and `data` values as passed to the function

```{r, eval=TRUE}
linreg_mod <- linreg$new(formula = Petal.Length ~ Species, data = iris)
```

The results of the model can be viewed using the implemented RC class methods as shown in the following examples:

### Method print()

Prints out the regression coefficients and coefficient names
```{r, eval=TRUE}
linreg_mod$print()
```

### Method plot()

Plots two graphs - Residuals vs Fitted graph and Scale - Location

```{r, eval=TRUE, warning=FALSE}
linreg_mod$plot()
```
```{r, eval=TRUE, warning=FALSE}
linreg_mod$plot()
```

### Method resid()

Returns the vector of residuals

```{r, eval=TRUE}
head(linreg_mod$resid())
```

### Method pred()

Returns the predicted values

```{r, eval=TRUE}
head(linreg_mod$pred())
```

### Method coef()

Returns the regression coefficients as a named vector
    
```{r, eval=TRUE}
linreg_mod$coef()
```

### Method summary()

Prints out the regression coefficients with their standard error, t-values and p-values, along with the estimate of residual standard error and degrees of freedom in the model

```{r, eval=TRUE}
linreg_mod$summary()
```
