---
title: "API Kolada"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{apikolada}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{httr, jsonlite}    
---

```{r, include = FALSE}
library(apikolada)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette describes the use of RC class **`api.kolada`** to interface with Kolada's web API. Kolada provides a web-service for accessing standardized key performance indicators (KPI) concerning Swedish municipalities and organizational units. More details about the API and its database can be found at the [kolada API web page](https://www.kolada.se/index.php?_p=index/API).

## Class api.kolada()

The implemented methods of the class can be used to access and analyze the KPI data of municipalities on kolada database.

In the examples shown below, `obj.kolada` is a class object of type api.kolada(). 
```{r, eval=TRUE}
obj.kolada <- api.kolada$new()
```

### Method get.muni()

Gets all the municipality codes from API.

```{r, eval=TRUE}
muni <- obj.kolada$get.muni()
head(muni)
```

### Method get.ou.muni(muni)

Gets all the organizational units in a municipality `muni` from API.

```{r, eval=TRUE}
ou <- obj.kolada$get.ou.muni("Stockholm")
head(ou)
```

### Method get.kpi.group()

Gets all KPI groups from API.

```{r, eval=TRUE}
kpi.group <- obj.kolada$get.kpi.group()
head(kpi.group[,-2])
```

### Method get.kpi.member(kpigroup)

Gets the KPI indicators within the selected KPI group `kpigroup`.

```{r, eval=TRUE}
kpi.members <- obj.kolada$get.kpi.member("Befolkning")
head(kpi.members)
```

### Method get.muni.kpi(kpigroup, muni, kpi, gender)

Gets the KPI data for a municipality `muni` based on the chosen KPI indicator `kpi` from API filtered on `gender`, if applicable.

```{r, eval=TRUE}
muni.kpi <- obj.kolada$get.muni.kpi("Befolkning", "Linköping", "Invånare 65-79 år, antal", "T")
head(muni.kpi)
```

### Method plot.muni.kpi(period_val, kpi_val, kpi_label)

Plots KPI values `kpi_val` of the municipality over the years `period_val` for the selected KPI indicator `kpi_label`.

```{r, eval=TRUE}
obj.kolada$plot.muni.kpi(muni.kpi$period, muni.kpi$kpi, "Invånare 65-79 år, antal")
```

### Method summary.kpi(kpi_df)

Gives the summary statistics of municipality KPI values `kpi_df`.

```{r, eval=TRUE}
obj.kolada$summary.kpi(muni.kpi)
```
